### 2018-06-25 doc/methods.tex
lots of plots looking at properties of the data
rewrote model statement. need to expand

Second, we considered two aspects of global temperature: mean temperature during interval \(t\), and the mean temperature during interval \(t - 1\).

Our model had up to three non-nested varying intercepts: age of observation \(a\), bin of observation \(b\), and taxonomic group of the observation \(c\). Age of observation is a critical varying intercept as this effect, when considered along with the intercept \(\beta_{0}\), captures changes to baseline hazard with species age. Because these factors are modeled as varying intercepts, we are assuming that the effects of geographic range (\(\beta_{1} \dots \beta_{3}\)) and temperature (\(\beta_{4} \dots \beta_{5}\)) are constant over time.

The logistic regression can thus be expressed as 
\begin{equation}
  \begin{aligned}
    t_{i} = logit^{-1}(\beta_{0} + \beta_{1} r_{i} + \beta_{2} d_{i} + \beta_{3} r_{i} d_{i} + a_{j[i]} + b_{k[i]} + c_{f[i]}) \\
  \end{aligned}
  \label{eq:core}
\end{equation}
with \(i\) indexing the observation and bracket subscripts referencing the class of the \(i\)th observation e.g. \(j[i]\) is the age of the \(i\)th observation. \(a_{j[i]}\) is taxon age in bins at time of observation where \(j = 1, 2, \cdots, J\) and \(J\) is the maximum observed age of any taxon. \(b_{k}\) is the time of observation where \(k = 1, 2, \cdots, K\) and \(K\) is the number of time bins observed. Finally, \(c_{f}\) is the taxonomic group of an observation where \(f = 1, 2, \cdots, F)\) and \(F\) is the total number of taxonomic groups analyzed.

>>>>

Second, we considered two aspects of global temperature: mean temperature during interval \(t\), and the lag of mean temperature (i.e. mean temperature during interval \(t - 1\).)

% lots of ways of varying
% maxgcd * diff_maxgcd varies first by when then by fossil group.
% non-nested varying intercept based on age, then by fossil group.

% event ~ temp + lag1_temp + maxgcd * diff_maxgcd + 
% (1 + maxgcd * diff_maxgcd | fact_mybin/fossil.group) +
% (1 | fact_relage/fossil.group)
The logistic regression can thus be expressed as 
% scripts
%   t = death? 0 if not last observation
%   \beta = effect range vars (varies over time and by fossil group)
%   \gamma = effect temp vars
%   a = age effect
% subscripts
%   w = when in My
%   f = fossil group
%   d = age at observation
% 
\begin{equation}
  \begin{aligned}
    t_{i} &= logit^{-1}(\beta_{w[i], f[i]} + \beta_{w[i], f[i]} range_{i} + \beta_{w[i], f[i]} rangediff_{i} + \beta_{w[i], f[i]} range_{i} rangediff_{i} + \gamma_{1} temp_{w[i]} + \gamma_{2} templag_{w[i]} + a_{d[i], f[i]}) \\
    \beta_{w, f} &\sim MVN(\alpha_{f}, \Sigma) \\
    \alpha_{f} &\sim MVN(0, \Sigma) \\
    \gamma &\sim N(0, \sigma) \\
    a_{w, f} &\sim MVN(\delta_{f}, \Sigma) \\
    \delta_{f} &\sim MVN(0, \Sigma) \\
  \end{aligned}
  \label{eq:core}
\end{equation}
with \(i\) indexing the observation and bracket subscripts referencing the class of the \(i\)th observation e.g. \(j[i]\) is the age of the \(i\)th observation. \(a_{j[i]}\) is taxon age in bins at time of observation where \(j = 1, 2, \cdots, J\) and \(J\) is the maximum observed age of any taxon. \(b_{k}\) is the time of observation where \(k = 1, 2, \cdots, K\) and \(K\) is the number of time bins observed. Finally, \(c_{f}\) is the taxonomic group of an observation where \(f = 1, 2, \cdots, F)\) and \(F\) is the total number of taxonomic groups analyzed.
